---
title: "Fungal Growth Analysis"
author: "Your Name"
output: 
  html_document:
    theme: united
    toc: true
    toc_float: true
    code_folding: hide
---

# Model Selection {#tab-model-selection}

In this section, we will compare the performance of three different models: **Linear**, **Exponential**, and **Log-Transformed** for each fungal species at various temperatures. We will compare the following metrics:

- **Residual Sum of Squares (RSS)**: Lower values indicate a better fit.
- **Adjusted R-squared**: Higher values indicate a better fit.
- **AIC (Akaike Information Criterion)**: Lower values indicate a better model fit considering model complexity.
- **BIC (Bayesian Information Criterion)**: Lower values indicate a better model fit, penalizing model complexity more heavily than AIC.

```{r load-packages, echo=FALSE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(viridis)
library(tidyr)
library(car)
library(multcompView)
library(agricolae)
library(ggsignif)
library(report)
library(multcomp)
library(growthcurver)
library (gridExtra)
library(caret)
library(kableExtra)
library(plotly)
library(ggtext)
library(glue)
library(ggtext)
library(patchwork) 
```

```{r}
df <- read.csv("Rosalynn Tidy data.csv")
#Filter week 1-5, Suillus data up to week 6. 
df_1_5 <- df %>%
  filter(Week %in% 1:5)
# Convert relevant columns to factors
df_1_5$Temperature..Celsius. <- as.factor(df_1_5$Temperature..Celsius.)
#df_1_5$Week <- as.factor(df_1_5$Week)
df_1_5$Species <- as.factor(df_1_5$Species)
# Rename the column
colnames(df_1_5)[colnames(df_1_5) == "Temperature..Celsius."] <- "Temperature"

# Initialize an empty list to store model metrics
model_metrics <- list()

# Loop over each species and temperature to fit models and calculate metrics
for(species in unique(df_1_5$Species)) {
  for(temp in unique(df_1_5$Temperature)) {
    
    # Filter the data for this species and temperature
    df_subset <- df_1_5 %>% filter(Species == species, Temperature == temp)
    
    # Fit the Linear Model
    linear_model <- lm(Growth ~ Week, data = df_subset)
    
    # Fit the Exponential Model (log-transformed linear)
    exp_model <- lm(log(Growth) ~ Week, data = df_subset)
    
    # Fit the Logarithmic Model (Growth ~ log(Week))
    log_fit_model <- lm(Growth ~ log(Week), data = df_subset)
    
    # Calculate Metrics for each model
    metrics <- data.frame(
      Species = species,
      Temperature = temp,
      Metric = c("RSS", "Adjusted R-squared", "AIC", "BIC"),
      Linear_Model = c(sum(residuals(linear_model)^2),
                       summary(linear_model)$adj.r.squared,
                       AIC(linear_model),
                       BIC(linear_model)),
      Exponential_Model = c(sum(residuals(exp_model)^2),
                            summary(exp_model)$adj.r.squared,
                            AIC(exp_model),
                            BIC(exp_model)),
      Logarithmic_Model = c(sum(residuals(log_fit_model)^2),
                            summary(log_fit_model)$adj.r.squared,
                            AIC(log_fit_model),
                            BIC(log_fit_model))
    )
    
    # Determine the best model for each metric
    metrics$Best_Model <- sapply(1:nrow(metrics), function(i) {
      if (metrics$Metric[i] == "Adjusted R-squared") {
        # For Adjusted R-squared, higher is better
        if (metrics$Linear_Model[i] == max(metrics$Linear_Model[i], metrics$Exponential_Model[i], metrics$Logarithmic_Model[i])) {
          "Linear Model"
        } else if (metrics$Exponential_Model[i] == max(metrics$Linear_Model[i], metrics$Exponential_Model[i], metrics$Logarithmic_Model[i])) {
          "Exponential Model"
        } else {
          "Logarithmic Model"
        }
      } else {
        # For RSS, AIC, and BIC, lower is better
        if (metrics$Linear_Model[i] == min(metrics$Linear_Model[i], metrics$Exponential_Model[i], metrics$Logarithmic_Model[i])) {
          "Linear Model"
        } else if (metrics$Exponential_Model[i] == min(metrics$Linear_Model[i], metrics$Exponential_Model[i], metrics$Logarithmic_Model[i])) {
          "Exponential Model"
        } else {
          "Logarithmic Model"
        }
      }
    })
    
    # Store the metrics in the list
    model_metrics[[paste(species, temp, sep = "_")]] <- metrics
  }
}

# Combine all metrics into one data frame
model_comparison <- do.call(rbind, model_metrics)
```

```{r}
# Display the table of model comparison with scrolling enabled for rows
model_comparison %>%
  kbl(caption = "Model Comparison for Different Species and Temperatures") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = TRUE, 
                fixed_thead = TRUE) %>%
  scroll_box(height = "400px")
```




# Model Building: Exponential Models {#tab-model-building}

In this section, we build **exponential models** to describe the growth of fungi at different temperatures. We will visualize both the **raw data** and the **predicted data** from the exponential models, and offer the option to compare them in the same plot.

```{r}
# Create a list to store the fitted models and predicted data
exponential_models <- list()
predicted_data_list <- list()

# Loop over each species and temperature to fit exponential models
for(species in unique(df_1_5$Species)) {
  for(temp in unique(df_1_5$Temperature)) {
    
    # Filter the data for this species and temperature
    df_subset <- df_1_5 %>% filter(Species == species, Temperature == temp)
    
    # Fit the Exponential Model (log-transformed linear)
    exp_model <- lm(log(Growth) ~ Week, data = df_subset)
    
    # Store the model
    model_name <- paste("Exp_Model", species, temp, sep = "_")
    exponential_models[[model_name]] <- exp_model
    
    # Generate predicted values
    predicted_log_growth <- predict(exp_model, newdata = df_subset)
    predicted_growth <- exp(predicted_log_growth)  # Back-transform to the original scale
    
    # Store predicted data in a data frame
    df_predicted <- df_subset %>%
      mutate(Predicted_Growth = predicted_growth)
    
    # Store the predicted data for later use
    predicted_data_list[[model_name]] <- df_predicted
  }
}

# Combine all predicted data into one data frame for easier use
all_predicted_data <- bind_rows(predicted_data_list)

# Print a sample of the predicted data
head(all_predicted_data)
```

### Compare Raw and Fitted Data
```{r graph-data, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=7}
# Create a function to abbreviate species names with italic Markdown syntax
abbreviate_species <- function(species) {
  species <- gsub("Cenococcum geophilum", "<i>C. geophilum</i>", species)
  species <- gsub("Laccaria bicolor", "<i>L. bicolor</i>", species)
  species <- gsub("Rhizopogon Evadens", "<i>R. evadens</i>", species)
  species <- gsub("Rhizopogon idahoensis", "<i>R. idahoensis</i>", species)
  species <- gsub("Suillus lakei", "<i>S. lakei</i>", species)
  return(species)
}

# Apply the abbreviation function to Species column
df_1_5$Species <- abbreviate_species(df_1_5$Species)
all_predicted_data$Species <- abbreviate_species(all_predicted_data$Species)

# Custom labeller for temperature, to directly label 15°C, 20°C, 25°C
temperature_labeller <- as_labeller(c(
  "15" = "15°C",
  "20" = "20°C",
  "25" = "25°C"
))

# Plot raw data
plot_raw <- ggplot(df_1_5, aes(x = Week, y = Growth, color = Species)) +
  geom_point() +
  facet_grid(Temperature ~ Species, labeller = labeller(Temperature = temperature_labeller)) +  # Temperature as rows, Species as columns
  labs(title = "Raw Growth Data", x = "Week", y = "Growth (cm^2)") +
  theme_minimal() +
  theme(
    strip.text.x = element_markdown(size = 12),        # Use element_markdown for italic facet labels
    strip.text.y = element_text(size = 12),            # Adjust temperature label size
    axis.text = element_text(size = 10),               # Adjust axis text size
    plot.margin = margin(2, 2, 2, 2, "cm"),            # Add margins to avoid text overlap
    legend.position = "none"                           # Remove redundant legend
  )

# Plot predicted data
plot_predicted <- ggplot(all_predicted_data, aes(x = Week, y = Predicted_Growth, color = Species)) +
  geom_line() +
  facet_grid(Temperature ~ Species, labeller = labeller(Temperature = temperature_labeller)) +  # Temperature as rows, Species as columns
  labs(title = "Predicted Growth Data (Exponential Model)", x = "Week", y = "Predicted Growth (cm^2)") +
  theme_minimal() +
  theme(
    strip.text.x = element_markdown(size = 12),        # Use element_markdown for italic facet labels
    strip.text.y = element_text(size = 12),            # Adjust temperature label size
    axis.text = element_text(size = 10),               # Adjust axis text size
    plot.margin = margin(2, 2, 2, 2, "cm"),            # Add margins to avoid text overlap
    legend.position = "none"                           # Remove redundant legend
  )

# Combine raw and predicted data in one plot
plot_combined <- ggplot() +
  geom_point(data = df_1_5, aes(x = Week, y = Growth, color = Species)) +
  geom_line(data = all_predicted_data, aes(x = Week, y = Predicted_Growth, color = Species)) +
  facet_grid(Temperature ~ Species, labeller = labeller(Temperature = temperature_labeller)) +  # Temperature as rows, Species as columns
  labs(title = "Raw and Predicted Growth Data", x = "Week", y = "Growth (cm^2)") +
  theme_minimal() +
  theme(
    strip.text.x = element_markdown(size = 12),        # Use element_markdown for italic facet labels
    strip.text.y = element_text(size = 12),            # Adjust temperature label size
    axis.text = element_text(size = 10),               # Adjust axis text size
    plot.margin = margin(2, 2, 2, 2, "cm"),            # Add margins to avoid text overlap
    legend.position = "none"                           # Remove redundant legend
  )

# Convert ggplot objects to interactive plots using plotly
plotly_raw <- ggplotly(plot_raw)
plotly_predicted <- ggplotly(plot_predicted)
plotly_combined <- ggplotly(plot_combined)

# Display the interactive plots
plotly_raw
plotly_predicted
plotly_combined
```
### Exponential Model Summary for Each Species at Each Temperature

The table below summarizes the results of fitting exponential models for each fungal species at different temperatures. For each combination of species and temperature, the following details are provided:

- **F-Value**: This is the F-statistic for the model, which assesses the overall significance of the model.
- **P-Value**: The P-value indicates the significance of the model. A lower P-value (< 0.05) suggests that the model is statistically significant.
- **Degrees of Freedom**: This represents the number of data points minus the number of parameters estimated by the model.
- **Model Equation**: The fitted equation for the model, represented as a linear relationship between the logarithm of growth and time (weeks).
```{r}
# Initialize a data frame to store the summary information
model_summary <- data.frame(Species = character(), 
                            Temperature = numeric(), 
                            F_Value = numeric(), 
                            P_Value = numeric(), 
                            Degrees_of_Freedom = character(), 
                            Equation = character(),
                            stringsAsFactors = FALSE)

# Loop over each species and temperature to fit exponential models and extract metrics
for(species in unique(df_1_5$Species)) {
  for(temp in unique(df_1_5$Temperature)) {
    
    # Filter the data for this species and temperature
    df_subset <- df_1_5 %>% filter(Species == species, Temperature == temp)
    
    # Fit the exponential model (log-transformed linear)
    exp_model <- lm(log(Growth) ~ Week, data = df_subset)
    
    # Extract model summary details
    model_summary_info <- summary(exp_model)
    
    # Extract F-value, P-value, and degrees of freedom
    f_value <- model_summary_info$fstatistic[1]
    p_value <- pf(f_value, model_summary_info$fstatistic[2], model_summary_info$fstatistic[3], lower.tail = FALSE)
    df1 <- model_summary_info$fstatistic[2]  # Numerator degrees of freedom
    df2 <- model_summary_info$fstatistic[3]  # Denominator degrees of freedom
    
    # Extract the model equation
    intercept <- coef(exp_model)[1]
    slope <- coef(exp_model)[2]
    equation <- paste0("log(Growth) = ", round(intercept, 4), " + ", round(slope, 4), " * Week")
    
    # Store the results in the data frame
    model_summary <- rbind(model_summary, data.frame(Species = species, 
                                                     Temperature = temp, 
                                                     F_Value = f_value, 
                                                     P_Value = p_value, 
                                                     Degrees_of_Freedom = paste(df1, df2, sep = ", "), 
                                                     Equation = equation))
  }
}

# Create a scrollable table
model_summary %>%
  kbl(caption = "Exponential Model Summary for Each Species at Each Temperature") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, fixed_thead = TRUE) %>%
  scroll_box(height = "400px")
```

### Interactive Exponential Growth Curves for Each Species at Each Temperature

The plot below displays the **perfect exponential growth curves** for each fungal species at different temperatures. These curves are based on the **exponential model** fitted to the data, using the equation:

\[
\text{Growth} = e^{(\text{intercept} + \text{slope} \times \text{Week})}
\]

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=7}
## Initialize an empty data frame to store predicted data for the curves
predicted_curve_data <- data.frame(Species = character(), 
                                   Temperature = numeric(), 
                                   Week = numeric(), 
                                   Predicted_Growth = numeric(), 
                                   stringsAsFactors = FALSE)

# Loop over each species and temperature to generate curves
for(species in unique(df_1_5$Species)) {
  for(temp in unique(df_1_5$Temperature)) {
    
    # Filter the data for this species and temperature
    df_subset <- df_1_5 %>% filter(Species == species, Temperature == temp)
    
    # Fit the exponential model (log-transformed linear)
    exp_model <- lm(log(Growth) ~ Week, data = df_subset)
    
    # Extract coefficients for the model (intercept and slope)
    intercept <- coef(exp_model)[1]
    slope <- coef(exp_model)[2]
    
    # Generate predicted values for a sequence of weeks (1 to 5, or based on your data)
    weeks <- seq(min(df_1_5$Week), max(df_1_5$Week), length.out = 100)  # 100 points for a smooth curve
    
    # Compute the predicted growth based on the model equation: Growth = exp(intercept + slope * Week)
    predicted_growth <- exp(intercept + slope * weeks)
    
    # Store the predicted data for this species and temperature
    predicted_curve_data <- rbind(predicted_curve_data, 
                                  data.frame(Species = species, 
                                             Temperature = temp, 
                                             Week = weeks, 
                                             Predicted_Growth = predicted_growth))
  }
}

# Custom labeller for temperature, to directly label 15°C, 20°C, 25°C
temperature_labeller <- as_labeller(c(
  "15" = "15°C",
  "20" = "20°C",
  "25" = "25°C"
))

# Plot the perfect curves for the exponential model
plot_perfect_curves <- ggplot(predicted_curve_data, aes(x = Week, y = Predicted_Growth, color = Species)) +
  geom_line(size = 1.2) +  # Use geom_line to draw smooth curves
  facet_grid(Temperature ~ Species, labeller = labeller(Temperature = temperature_labeller)) +  # Temperature as rows, Species as columns
  labs(title = "Perfect Exponential Growth Curves Based on Model", x = "Week", y = "Predicted Growth (cm^2)") +
  theme_minimal() +
  theme(
    strip.text.x = element_markdown(size = 12),        # Use element_markdown for italic facet labels
    strip.text.y = element_text(size = 12),            # Adjust temperature label size
    axis.text = element_text(size = 10),               # Adjust axis text size
    plot.margin = margin(2, 2, 2, 2, "cm"),            # Add margins to avoid text overlap
    legend.position = "none"                           # Remove redundant legend
  )

# Convert the ggplot object to an interactive plotly object
interactive_plot <- ggplotly(plot_perfect_curves)

# Display the interactive plot
interactive_plot
```

### Calculation of the Derivative Function for the Exponential Model

For each fungal species at each temperature, an **exponential growth model** was fitted to the data. The general form of the model is:

\[
\text{Growth} = e^{(\text{intercept} + \text{slope} \times \text{Week})}
\]

This model describes how the growth of the fungi changes over time. To determine how fast the growth changes with respect to time (week), we need to compute the **derivative** of the growth function with respect to the week.

#### Step 1: Apply the Chain Rule
To compute the derivative of the growth function, we apply the **chain rule**. The chain rule tells us how to differentiate composite functions like this one, where we have an exponent that depends on another variable (week). 

We begin by differentiating the exponential function:

\[
\frac{d}{d\text{Week}} \left( e^{(\text{intercept} + \text{slope} \times \text{Week})} \right)
\]

The chain rule gives us two parts:
1. The derivative of the exponential function, which is the same exponential function:
   \[
   e^{(\text{intercept} + \text{slope} \times \text{Week})}
   \]
2. The derivative of the exponent \((\text{intercept} + \text{slope} \times \text{Week})\) with respect to **week**. The derivative of this linear term is just the **slope**:
   \[
   \frac{d}{d\text{Week}} (\text{intercept} + \text{slope} \times \text{Week}) = \text{slope}
   \]

#### Step 2: Combine the Results
Multiplying the results of these two derivatives (following the chain rule), we get:

\[
\frac{d(\text{Growth})}{d(\text{Week})} = \text{slope} \times e^{(\text{intercept} + \text{slope} \times \text{Week})}
\]

#### Step 3: Substitution
We know from the original model that:

\[
\text{Growth} = e^{(\text{intercept} + \text{slope} \times \text{Week})}
\]

So, we can substitute this back into the derivative equation. Replacing the exponential term with **Growth**, we arrive at the simplified form:

\[
\frac{d(\text{Growth})}{d(\text{Week})} = \text{Growth} \times \text{slope}
\]

#### What the Derivative Tells Us:
- The derivative function tells us the **rate of growth** at any given point in time (week).
- It shows that the rate of growth is proportional to the current **growth** value, scaled by the **slope** of the model.
- As the fungi grow larger (as Growth increases), their growth rate also increases, but it is controlled by the value of the slope. 

In the table below, for each species and temperature combination, the derivative is expressed in the simplified form:

\[
\frac{d(\text{Growth})}{d(\text{Week})} = \text{Growth} \times \text{slope}
\]

This equation provides a clear understanding of how the rate of growth evolves over time for each species at each temperature.


```{r derivative equation calculation}
# Initialize a data frame to store the simplified derivative functions
derivative_summary <- data.frame(Species = character(), 
                                 Temperature = numeric(), 
                                 Derivative_Function = character(),
                                 stringsAsFactors = FALSE)

# Loop over each species and temperature to compute simplified derivative functions
for(species in unique(df_1_5$Species)) {
  for(temp in unique(df_1_5$Temperature)) {
    
    # Filter the data for this species and temperature
    df_subset <- df_1_5 %>% filter(Species == species, Temperature == temp)
    
    # Fit the exponential model (log-transformed linear)
    exp_model <- lm(log(Growth) ~ Week, data = df_subset)
    
    # Extract the slope of the model
    slope <- coef(exp_model)[2]
    
    # The simplified derivative function is Growth * slope
    derivative_function <- paste0("d(Growth)/d(Week) = Growth * ", round(slope, 4))
    
    # Store the derivative function for this species and temperature
    derivative_summary <- rbind(derivative_summary, 
                                data.frame(Species = species, 
                                           Temperature = temp, 
                                           Derivative_Function = derivative_function))
  }
}

# Display the final table
library(knitr)
library(kableExtra)

# Create a scrollable table
derivative_summary %>%
  kbl(caption = "Simplified Derivative Functions of Exponential Models for Each Species at Each Temperature") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, fixed_thead = TRUE) %>%
  scroll_box(height = "400px")

```

### Rate of Growth Curves Based on the Derivative of the Exponential Model
```{r derivative graphs, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=7}
# Initialize an empty data frame to store the derivative curve data
derivative_curve_data <- data.frame(Species = character(), 
                                    Temperature = numeric(), 
                                    Week = numeric(), 
                                    Rate_of_Growth = numeric(), 
                                    stringsAsFactors = FALSE)

# Loop over each species and temperature to compute derivative curves
for(species in unique(df_1_5$Species)) {
  for(temp in unique(df_1_5$Temperature)) {
    
    # Filter the data for this species and temperature
    df_subset <- df_1_5 %>% filter(Species == species, Temperature == temp)
    
    # Fit the exponential model (log-transformed linear)
    exp_model <- lm(log(Growth) ~ Week, data = df_subset)
    
    # Extract coefficients for the model (intercept and slope)
    intercept <- coef(exp_model)[1]
    slope <- coef(exp_model)[2]
    
    # Generate predicted growth for a sequence of weeks (1 to max week from the data)
    weeks <- seq(min(df_1_5$Week), max(df_1_5$Week), length.out = 100)  # 100 points for smooth curve
    predicted_growth <- exp(intercept + slope * weeks)
    
    # Compute the rate of growth using the derivative equation: Rate = Growth * slope
    rate_of_growth <- predicted_growth * slope
    
    # Store the derivative (rate of growth) data for this species and temperature
    derivative_curve_data <- rbind(derivative_curve_data, 
                                   data.frame(Species = species, 
                                              Temperature = temp, 
                                              Week = weeks, 
                                              Rate_of_Growth = rate_of_growth))
  }
}

# Custom labeller for temperature, to directly label 15°C, 20°C, 25°C
temperature_labeller <- as_labeller(c(
  "15" = "15°C",
  "20" = "20°C",
  "25" = "25°C"
))

# Plot the derivative (rate of growth) curves
plot_derivative_curves <- ggplot(derivative_curve_data, aes(x = Week, y = Rate_of_Growth, color = Species)) +
  geom_line(size = 1.2) +  # Use geom_line to draw smooth curves
  facet_grid(Temperature ~ Species, labeller = labeller(Temperature = temperature_labeller)) +  # Temperature as rows, Species as columns
  labs(title = "Rate of Growth Curves Based on Derivative of Exponential Model", 
       x = "Week", y = "Rate of Growth (cm^2 per Week)") +
  theme_minimal() +
  theme(
    strip.text.x = element_markdown(size = 12),        # Use element_markdown for italic facet labels
    strip.text.y = element_text(size = 12),            # Adjust temperature label size
    axis.text = element_text(size = 10),               # Adjust axis text size
    plot.margin = margin(2, 2, 2, 2, "cm"),            # Add margins to avoid text overlap
    legend.position = "none"                           # Remove redundant legend
  )

# Convert the ggplot object to an interactive plotly object
interactive_derivative_plot <- ggplotly(plot_derivative_curves)

# Display the interactive plot
interactive_derivative_plot
```

### Raw Growth Rate

The **raw growth rate** was calculated for each fungal strain on each plate, at every temperature and time point (week). This method provides the growth rate for each individual replicate, without aggregating across plates. The growth rate measures how much the fungal colony's area increased between consecutive time points.

The growth rate was computed using the following formula:

\[
\text{Growth Rate} = \frac{\text{Growth at Week}_n - \text{Growth at Week}_{n-1}}{\text{Week}_n - \text{Week}_{n-1}}
\]
```{r raw growth rate, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=9 }

# Step 1: Calculate raw growth rate based on strain_ID
df_growth_rate <- df_1_5 %>%
  group_by(strain_ID, Temperature, Plate) %>%
  arrange(strain_ID, Temperature, Plate, Week) %>%
  mutate(
    Previous_Week = lag(Week),  # Previous week for each plate
    Previous_Growth = lag(Growth),  # Previous growth value
    Growth_Rate = ifelse(is.na(Previous_Week) | is.na(Previous_Growth), 0, (Growth - Previous_Growth) / (Week - Previous_Week))  # Growth rate calculation
  ) %>%
  ungroup()

# Custom labeller for temperature, to directly label 15°C, 20°C, 25°C
temperature_labeller <- as_labeller(c(
  "15" = "15°C",
  "20" = "20°C",
  "25" = "25°C"
))

# Step 2: Create a static ggplot to graph the growth rate by Species, with color based on Species
plot_growth_rate <- ggplot(df_growth_rate, aes(x = Week, y = Growth_Rate, color = Species)) +
  geom_point(size = 2) +  # Use geom_point to plot individual data points
  facet_grid(Temperature ~ Species, labeller = labeller(Temperature = temperature_labeller), scales = "free") +  # Facet by species and temperature with custom labeller
  labs(title = "Growth Rate for Each Species at Different Temperatures",
       x = "Week", y = "Growth Rate (cm^2 per Week)") +
  theme_minimal() +
  theme(
    strip.text.x = element_text(size = 10),  # Adjust species label size
    strip.text.y = element_text(size = 10),  # Adjust temperature label size
    axis.text = element_text(size = 8),      # Adjust axis text size
    plot.margin = margin(2, 2, 2, 2, "cm"),  # Add margins
    legend.position = "none"                 # Eliminate the legend
  )

# Step 3: Convert the static plot to an interactive plot using plotly
interactive_growth_rate_plot <- ggplotly(plot_growth_rate)

# Step 4: Display the interactive plot
interactive_growth_rate_plot
```
###Aggregated Raw Growth Rate for Each Species at Different Temperatures

This figure displays the **aggregated raw growth rates** for multiple fungal species grown at various temperatures (15°C, 20°C, and 25°C). The **growth rate** is calculated based on the change in the **mean growth** between consecutive weeks, aggregated across multiple replicate plates for each species at each temperature.

```{r aggregated growth rate, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=7 }
# Step 1: Aggregate the data by calculating the mean growth for each species, temperature, and week across plates
df_aggregated <- df_1_5 %>%
  group_by(Species, Temperature, Week) %>%
  summarize(
    Mean_Growth = mean(Growth, na.rm = TRUE)  # Calculate mean growth across plates
  ) %>%
  arrange(Species, Temperature, Week)

# Step 2: Calculate the growth rate using the aggregated data
df_aggregated_growth_rate <- df_aggregated %>%
  mutate(
    Previous_Week = lag(Week),  # Previous week for each species and temperature
    Previous_Mean_Growth = lag(Mean_Growth),  # Previous mean growth value
    Growth_Rate = ifelse(is.na(Previous_Week) | is.na(Previous_Mean_Growth), 0, 
                         (Mean_Growth - Previous_Mean_Growth) / (Week - Previous_Week))  # Calculate mean growth rate
  ) %>%
  ungroup()

# Custom labeller for temperature, to directly label 15°C, 20°C, 25°C
temperature_labeller <- as_labeller(c(
  "15" = "15°C",
  "20" = "20°C",
  "25" = "25°C"
))

# Step 3: Create the plot for aggregated raw growth rate
plot_aggregated_growth_rate <- ggplot(df_aggregated_growth_rate, aes(x = Week, y = Growth_Rate, color = Species)) +
  geom_point(size = 2) +  # Plot aggregated growth rate points
  geom_line()+
  facet_grid(Temperature ~ Species, labeller = labeller(Temperature = temperature_labeller), scales = "free") +  # Facet by species and temperature
  labs(title = "Aggregated Raw Growth Rate for Each Species at Different Temperatures",
       x = "Week", y = "Mean Growth Rate (cm^2 per Week)") +
  theme_minimal() +
  theme(
    strip.text.x = element_text(size = 10),  # Adjust species label size
    strip.text.y = element_text(size = 10),  # Adjust temperature label size
    axis.text = element_text(size = 8),      # Adjust axis text size
    plot.margin = margin(2, 2, 2, 2, "cm"),  # Add margins
    legend.position = "none"                 # Remove legend
  )

# Step 4: Convert to an interactive plot using plotly (optional)
interactive_aggregated_growth_rate_plot <- ggplotly(plot_aggregated_growth_rate)

# Step 5: Display the interactive plot
interactive_aggregated_growth_rate_plot
```


### Generated Niche Curve

This niche curve represents the **growth rate at week 5** (The highest growth rate) for various fungal species grown at different temperatures (15°C, 20°C, and 25°C). The **growth rate** was calculated using the derivative of an exponential growth model fitted to each species' growth data.
```{r model niche curve,message=FALSE, warning=FALSE}
# Step 1: Fit an exponential model for each species and temperature, and calculate the growth rate at Week 5
df_growth_rate_week5 <- data.frame(Species = character(), 
                                   Temperature = numeric(), 
                                   Growth_Rate_Week5 = numeric(), 
                                   stringsAsFactors = FALSE)

for(species in unique(df_1_5$Species)) {
  for(temp in unique(df_1_5$Temperature)) {
    
    # Filter the data for this species and temperature
    df_subset <- df_1_5 %>% filter(Species == species, Temperature == temp)
    
    # Fit the exponential model (log-transformed linear)
    exp_model <- lm(log(Growth) ~ Week, data = df_subset)
    
    # Extract coefficients for the model (intercept and slope)
    intercept <- coef(exp_model)[1]
    slope <- coef(exp_model)[2]
    
    # Calculate the growth at Week 5
    growth_at_week5 <- exp(intercept + slope * 5)  # Growth at week 5
    
    # Calculate the derivative (growth rate) at Week 5
    growth_rate_at_week5 <- growth_at_week5 * slope
    
    # Store the results in the dataframe
    df_growth_rate_week5 <- rbind(df_growth_rate_week5, 
                                  data.frame(Species = species, 
                                             Temperature = as.numeric(temp),  # Ensure Temperature is numeric
                                             Growth_Rate_Week5 = growth_rate_at_week5))
  }
}

# Step 2: Create the plot of maximum growth rate at week 5 for each temperature
plot_growth_rate_temp <- ggplot(df_growth_rate_week5, aes(x = Temperature, y = Growth_Rate_Week5, color = Species)) +
  geom_point(size = 4) +  # Plot points for growth rate at week 5
  geom_line(aes(group = Species), size = 1.2) +  # Connect points by species
  scale_x_continuous(breaks = c(15, 20, 25), labels = c("15°C", "20°C", "25°C")) +  # Label temperature in °C
  labs(title = "Niche Curve of Different Fungi",
       x = "Temperature (°C)", y = "Growth Rate at Week 5 (cm^2 per Week)") +
  theme_minimal() +
  theme(
    strip.text.x = element_text(size = 10),  # Adjust species label size
    axis.text = element_text(size = 10),     # Adjust axis text size
    axis.title = element_text(size = 12),    # Adjust axis title size
    plot.title = element_text(size = 14, face = "bold"),  # Adjust title size
    legend.position = "right"                # Keep the legend for species
  )

# Step 3: Convert to interactive plot using plotly
interactive_plot_growth_rate_temp <- ggplotly(plot_growth_rate_temp)

# Step 4: Display the interactive plot
interactive_plot_growth_rate_temp
```

### Raw Aggregated Niche Curve
```{r Raw Niche curve, message=FALSE, warning=FALSE}
# Step 1: Aggregate the data by calculating the mean growth for each species, temperature, and week across plates
df_aggregated <- df_1_5 %>%
  group_by(Species, Temperature, Week) %>%
  summarize(
    Mean_Growth = mean(Growth, na.rm = TRUE)  # Calculate mean growth across plates
  ) %>%
  arrange(Species, Temperature, Week)

# Step 2: Calculate the growth rate using the aggregated data
df_aggregated_growth_rate <- df_aggregated %>%
  mutate(
    Previous_Week = lag(Week),  # Previous week for each species and temperature
    Previous_Mean_Growth = lag(Mean_Growth),  # Previous mean growth value
    Growth_Rate = ifelse(is.na(Previous_Week) | is.na(Previous_Mean_Growth), 0, 
                         (Mean_Growth - Previous_Mean_Growth) / (Week - Previous_Week))  # Calculate mean growth rate
  ) %>%
  ungroup()

# Step 3: Find the maximum growth rate for each species and temperature
df_max_growth_rate <- df_aggregated_growth_rate %>%
  group_by(Species, Temperature) %>%
  summarize(Max_Growth_Rate = max(Growth_Rate, na.rm = TRUE)) %>%
  mutate(Temperature = as.numeric(Temperature)) %>%  # Convert Temperature to numeric
  ungroup()

# Step 4: Create the plot of maximum growth rate for each temperature
plot_max_growth_rate_temp <- ggplot(df_max_growth_rate, aes(x = Temperature, y = Max_Growth_Rate, color = Species)) +
  geom_point(size = 4) +  # Plot points for max growth rate
  geom_line(aes(group = Species), size = 1.2) +  # Connect points by species
  scale_x_continuous(breaks = c(15, 20, 25), labels = c("15°C", "20°C", "25°C")) +  # Label temperature in °C
  labs(title = "Raw Niche Curve",
       x = "Temperature (°C)", y = "Maximum Growth Rate (cm^2 per Week)") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),     # Adjust axis text size
    axis.title = element_text(size = 12),    # Adjust axis title size
    plot.title = element_text(size = 14, face = "bold"),  # Adjust title size
    legend.position = "right"                # Keep the legend for species
  )

# Step 5: Convert to interactive plot using plotly
interactive_plot_max_growth_rate_temp <- ggplotly(plot_max_growth_rate_temp)

# Step 6: Display the interactive plot
interactive_plot_max_growth_rate_temp
```



